\begin{cv}{}
\begin{cvlist}{Education}
\item[Since 04/2018] \textbf{Dissertation}\\
Technical University Munich
\item[09/2014--03/2018] \textbf{Dissertation}\\
Gottfried Willhelm Leibniz Universit\"at Hannover
\item[10/2012--07/2014] \textbf{Master of Science}\\
Gottfried Willhelm Leibniz Universit\"at Hannover
\item[10/2007--09/2012] \textbf{Bachelor of Science}\\
Gottfried Willhelm Leibniz Universit\"at Hannover
\item[08/2003--07/2007] \textbf{High School Diploma}\\
Fachgymnasium Wirtschaft (Wunstorf)
\end{cvlist}

\begin{cvlist}{Professional Experience}
\item[Since 04/2022] \textbf{Franka Emika GmbH}\\
Head of Robotics Learning
\item[04/2018--03/2022] \textbf{Research Assistant}\\
Technical University Munich
\item[09/2014-03/2018] \textbf{Research Assistant}\\
Gottfried Willhelm Leibniz Universit\"at Hannover
\end{cvlist}

\begin{cvlist}{KPI Statistics}
    \item[Citations] $512$
    \item[h-index] $9$
    \item[i10-index] $7$
\end{cvlist}

\end{cv}

\subsection*{Publications}

\subsubsection*{Conference Papers}
\begin{itemize}
    \item \textbf{Johannsmeier, L.}, \& Haddadin, S. (2022, October). Can we reach human expert programming performance? A tactile manipulation case study in learning time and task performance. In Proc. International Conference on Intelligent Robots and Systems (IROS) (pp. 12081-12088). IEEE.
    \item Chen, X., \textbf{Johannsmeier, L.}, Sadeghian, H., Shahriari, E., Danneberg, M., Nicklas, A., ... \& Haddadin, S. (2022, October). On the Communication Channel in Bilateral Teleoperation: An Experimental Study for Ethernet, WiFi, LTE and 5G. In Proc. International Conference on Intelligent Robots and Systems (IROS) (pp. 7712-7719). IEEE.
    \item Voigt, F., \textbf{Johannsmeier, L.}, \& Haddadin, S. (2020). Multi-Level Structure vs. End-to-End-Learning in High-Performance Tactile Robotic Manipulation. In Proc. Conference on Robot Learning (CoRL) (pp. 2306-2316).
    \item Grischke, J., \textbf{Johannsmeier, L.}, Eich, L., \& Haddadin, S. (2019, May). Dentronics: review, first concepts and pilot study of a new application domain for collaborative robots in dental assistance. In Proc. International Conference on Robotics and Automation (ICRA) (pp. 6525-6532). IEEE.
    \item Kuhn, J., Ringwald, J., Schappler, M., Johannsmeier, L., \& Haddadin, S. (2019, May). Towards semi-autonomous and soft-robotics enabled upper-limb exoprosthetics: first concepts and robot-based emulation prototype. In Proc. International Conference on Robotics and Automation (ICRA) (pp. 9180-9186). IEEE.
    \item \textbf{Johannsmeier, L.}, Gerchow, M., \& Haddadin, S. (2019, May). A framework for robot manipulation: Skill formalism, meta learning and adaptive control. In Proc. International Conference on Robotics and Automation (ICRA) (pp. 5844-5850). IEEE.
    \item Haddadin, S., \& Johannsmeier, L. (2018, October). The art of manipulation: Learning to manipulate blindly. In Proc. International Conference on Intelligent Robots and Systems (IROS) (pp. 1-9). IEEE.
    \item Grassmann, R., \textbf{Johannsmeier, L.}, \& Haddadin, S. (2018, October). Smooth Point-to-Point Trajectory Planning in $ SE $(3) with Self-Collision and Joint Constraints Avoidance. In Proc. International Conference on Intelligent Robots and Systems (IROS) (pp. 1-9). IEEE.
    \item Shahriari, E., \textbf{Johannsmeier, L.}, \& Haddadin, S. (2018, June). Valve-based virtual energy tanks: A framework to simultaneously passify controls and embed control objectives. In Proc. American Control Conference (ACC) (pp. 3634-3641). IEEE.
    \item Haddadin, S., \textbf{Johannsmeier, L.}, Becker, M., Schappler, M., Lilge, T., Haddadin, S., ... \& Parusel, S. (2018, March). roboterfabrik: a pilot to link and unify German robotics education to match industrial and societal demands. In Proc. International Conference on Human-Robot Interaction (pp. 375-375). IEEE.
\end{itemize}

\subsubsection*{Journals}
\begin{itemize}
    \item Ringwald, J., Schneider, S., Chen, L., Knobbe, D., \textbf{Johannsmeier, L.}, Swikir, A., \& Haddadin, S. (2023). Towards Task-Specific Modular Gripper Fingers: Automatic Production of Fingertip Mechanics. Robotics and Automation Letters (R-AL), 8(3), 1866-1873.
    \item Haddadin, S., Parusel, S., \textbf{Johannsmeier, L.}, Golz, S., Gabl, S., Walch, F., ... \& Haddadin, S. (2022). The franka emika robot: A reference platform for robotics research and education. IEEE Robotics \& Automation Magazine (RAM), 29(2), 46-64.
    \item Naceri, A., Elsner, J., Tr\"obinger, M., Sadeghian, H., \textbf{Johannsmeier, L.}, Voigt, F., ... \& Haddadin, S. (2022). Tactile Robotic Telemedicine for Safe Remote Diagnostics in Times of Corona: System Design, Feasibility and Usability Study. Robotics and Automation Letters (R-AL), 7(4), 10296-10303.
    \item Grischke, J., \textbf{Johannsmeier, L.}, Eich, L., Griga, L., \& Haddadin, S. (2020). Dentronics: Towards robotics and artificial intelligence in dentistry. Dental Materials, 36(6), 765-778.
    \item Shahriari, E., \textbf{Johannsmeier, L.}, Jensen, E., \& Haddadin, S. (2019). Power flow regulation, adaptation, and learning for intrinsically robust virtual energy tanks. Robotics and Automation Letters (R-AL), 5(1), 211-218.
    \item Haddadin, S., \textbf{Johannsmeier, L.}, \& Ledezma, F. D. (2018). Tactile robots as a central embodiment of the tactile internet. Proceedings of the IEEE, 107(2), 471-487.
    \item \textbf{Johannsmeier, L.}, \& Haddadin, S. (2016). A hierarchical human-robot interaction-planning framework for task allocation in collaborative industrial assembly processes. Robotics and Automation Letters (R-AL), 2(1), 41-48.
\end{itemize}

\subsubsection*{Workshops}
\begin{itemize}
    \item \textbf{Johannsmeier, L.}, Ringwald, J., Kuehn, J., \& Haddadin, S. (2016). Teleoperated semi-autonomous control of the lwr and a humanoid hand via the myo armband. Workshop in International Conference on Robotics and Automation (ICRA), 12, 13.
    \item \textbf{Johannsmeier, L.}, \& Haddadin, S. (2015). A framework for task allocation in collaborative industrial assembly processes. In 8th International Workshop on Human-Friendly Robotics.
\end{itemize}

\subsubsection*{Book Chapters}
\begin{itemize}
    \item A\"smann, U., Chen, L., Ebert, S., G\"ohringer, D., Grzelak, D., Hidalgo, D., ..., \textbf{Johannsmeier, L.}, ... \& Podlubne, A. (2021). Human–robot cohabitation in industry. In Tactile Internet (pp. 41-73). Academic Press.
\end{itemize}

\subsection*{Submission in Preparation}

\begin{itemize}
    \item \textbf{Johannsmeier, L.}, Schneider, S., Voigt, F., \&Haddadin, S. (2023). Motor Memory for Few-Shot Learning in Robotic Manipulation. Submitted to Science Robotics.
    \item \textbf{Johannsmeier, L.}, Schneider, S., Li, Y., Burdet, E., \&Haddadin, S. (2023). Robot Curricula: Process-Centric Industrial Manipulation Taxonomy for Organisation, Classification and Synthesis of Tactile Robot Skills. Submitted to Nature Machine Intelligence.
\end{itemize}